Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-29 paper/synthesis.tex
WIP

\section{Random Generators Prediction and Synthesis} \label{sec:synthesis}

Characterizing the distribution of values of an arbitrary random generator is a
hard task.
%
It requires to model every random choice that it could possibly make to generate
a value.
%
In a recent work \tocite, we have shown that it is possible to
\emph{analitically} predict the average distribution of data constructors
produced by a random generators automatically derived considering only ADT
definitions---like the one presented on Section \ref{sec:randomtesting}.
%
For this purpose, we found that random generation of ADT values can be
characterized using the theory of \emph{branching processes} \tocite.
%
This probabilistic theory was originally conceived to predict the growth and
extinction of royal family trees the Victorian Era, being later applied to a
wide variety of research areas.
%
In this work we extend this model to predict the average distribution of values
of random generators derived considering structural information coming from
pattern matchings and abstract interfaces.

Essentially, a branching process is a special kind of Markov process that models
the evolution of a population of \emph{individuals} of different kinds across
discrete time steps known as \emph{generations}.
%
Each kind of individual is expected to produce an average number of offspring of
(possibly) different kinds from one generation to the next one.
%
In our original setting, we shown that each different data constructor can be
considered as an individual on its own.
%
On the other hand, any ADT value can be seen as a tree where each node
represents a root data constructor and has its sub-expressions as
sub-trees---hence note the similarity with family trees.
%
In this light, each tree level of a generated value can be seen as a generation
of our generation model.

We note $G_n$ to the vector of real numbers that represents the average
distribution of constructors that a random generator produces in the $n$-th
generation, where each element of this vector represents each different data
constructor involved in the generation process.
%
Then, it is possible to predict expected number of constructors a the generation
$n$ by the means of the next formula:
%
\begin{align*}
  E[G_n]^T = E[G_0]^T \cdot M^n
\end{align*}

\begin{align*}
  E \left[ P_n \right]
  =
  \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
\end{align*}

\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n+1}
\end{align*}

\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  M_{T} = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

\begin{equation*}
  \beta_{Html} = \qquad \qquad
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      0 \\ % Text
      0 \\ % Sing
      1 \\ % Tag
      2 \\ % Join
      0 \\ % simplify#1
      2 \\ % simplify#1
      0 \\ % hr
      1 \\ % div
      1 \\ % bold
    };
    \draw[densely dashed] ([xshift=-3pt]v-4-1.south west) -- ([xshift=3pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
  \end{tikzpicture}
  \hspace{20pt}
  \mathcal{P}_{Html} =
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      p_{Text} \\
      p_{Single} \\
      p_{Tag}  \\
      p_{Join} \\
      p_{simplify\#1} \\
      p_{simplify\#2} \\
      p_{hr} \\
      p_{div} \\
      p_{bold} \\
    };
    \draw[densely dashed] ([xshift=-15pt]v-4-1.south west) -- ([xshift=15pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
  \end{tikzpicture}
\end{equation*}

@@

It is possible to provide further ``end-to-end'' guarantees on the generated
values if we restrict the generation process to only bare data constructors and
pattern matching values.
%
In particular, we can stop considering pattern matching values as atomic
constructions and start seeing them as compositions of data constructors.
%
This way is possible to obtain the average ``total'' number of data constructors
that our generators will produce, regardless of if they are generated on their
own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$.

The prediction only needs to consider the prediction of each data constructor
and add the average number of constructors that we will additionally generate
within pattern matching values.
%
For instance, we can calculate the total number of constructors |Text| and
|Join| that we will generate by expanding the average number of pattern values
$simplify\#1$ and $simplify\#2$ that we generate on average as follows:
%
{\small
  \begin{align*}
    E^\downarrow\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\! E\left[ simplify\#1 \right]
    + E\left[ simplify\#2 \right] \\
    E^\downarrow\!\left[ Join \right]
    = E\left[ Join \right]
    + E\left[ simplify\#1 \right]
    + 2\! \cdot\! E\left[ simplify\#2 \right]
  \end{align*}
}
%
Where $simplify\#1$ carries two |Text| and one |Join| data constructors---and
viceversa for $simplify\#2$.

>>>>

\section{Prediction of Distributions} \label{sec:synthesis}

Characterizing the distribution of values of an arbitrary random generator is a
hard task.
%
It requires to model every random choice that it could possibly make to generate
a value.
%
In a recent work \tocite, we have shown that it is possible to
\emph{analitically} predict the average distribution of data constructors
produced by a random generators automatically derived considering only ADT
definitions---like the one presented on Section \ref{sec:randomtesting}.
%
For this purpose, we found that random generation of ADT values can be
characterized using the theory of \emph{branching processes} \tocite.
%
This probabilistic theory was originally conceived to predict the growth and
extinction of royal family trees the Victorian Era, being later applied to a
wide variety of research areas.
%
In this work we extend this model to predict the average distribution of values
of random generators derived considering structural information coming from
functions pattern matchings and abstract interfaces.

Essentially, a branching process is a special kind of Markov process that models
the evolution of a population of \emph{individuals of different kinds} across
discrete time steps known as \emph{generations}.
%
Each kind of individual is expected to produce an average number of offspring of
(possibly) different kinds from one generation to the next one.
%
In our original setting, we shown that each different data constructor can be
considered as an individual on its own.
%
On the other hand, any ADT value can be seen as a tree where each node
represents a root data constructor and has its sub-expressions as
sub-trees---hence note the similarity with family trees.
%
In this light, each tree level of a random value can be seen as a generation of
individuals in this model.

Then, we can characterize the distribution of constructors that a random
generator produces in the $n$-th generaton as a vector $G_n$ that groups the
number of constructors of each kind produced in that generation.
We note $G_n$ to the vector of real numbers that represents the distribution of
%
Then, the we can predict the expected distribution of constructors at the
generation $n$ (noted as $E[G_n]$) using the following characteristic equation
from branching processes theory:
%
\begin{align}
  E[G_n]^T = E[G_0]^T \cdot M^n
  \label{eqn:prediction}
\end{align}

Where $E[G_0]$ represents the initial distribuition of constructors that our
generator produces, which simply consists of the generation probability of each
one.
%
The interesting aspect of the prediction mechanism is encoded in the matrix $M$,
known as a the \emph{mean matrix} of this stochastic process.
%
$M$ is a squared matrix with as rows and columns as different data constructors
involved in the generation process.
%
Each element $M_{i,j}$ of this matrix encodes the average number of data
constructors of kind $j$ that gets generated in a given level, provided that we
generated a constructor of kind $i$ at the previous level.
%
Is in this sense that this matrix encodes the ``branching'' behavior of our
random generation from one generation to the next one.
%
In our previous work, we show how each element of this matrix can be calculated
by exploiting type information from data type definitions.
%
For instance, the average number of |Text| data constructors that we will
generate provided that we generated a |Join| data constructor on the previous
level (corresponding to the element $M_{Join, Text}$) results $2\cdot p_{Text}$,
where $2$ is the number of the recursive fields of type |Html| present on the
constructor |Join| and $p_{Text}$ is the probability of generating the
constructor |Text|.
%
This pattern can be used to build the rest of the mean matrix as well.

\todo[inline]{Complete this!}

% \begin{align*}
%   E \left[ P_n \right]
%   = \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
% \end{align*}

Our realization in this work is that we can also consider that each different
pattern matching value and abstract interface function call can be considered as
another kind of individual belonging to the random generation process.
%
In this light, we can extend our mean matrix with this information, adding a row
and a column for each different pattern matching value and function call that we
add to the generation process, as we illustrate as follows:

\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  \hspace{20pt}
  M = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

To extend $M$ to consider the new possible constructions, we need to calculate
$M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
and $j$, respectively.
%
Fortunately, it is possible to follow the same pattern as we did on our original
work, this time considering the amount of recursive pattern variables if it the
case of a pattern matching value being a father individual, and the amount of
recursive arguments if we deal with a abstract interface function call as a
father individual.
%
For instance, the average amount of function calls to |div| that we will produce
provided that we generated a value matching the second pattern of |simplify| as a
father results $2\cdot p_{div}$.
%
Similarly, the average amount of values satisfying the first pattern matching of
|simplify| provided that we generated a |Tag| constructor as a father results
$1 \cdot p_{simplify\# 1}$.

Furthermore, we found that the prediction process can be factored in terms of
two vectors $\beta$ and $\mathcal{P}$, such that $\beta$ represents the number
of recursive sub-terms of each construction that we generate, whereas
$\mathcal{P}$ simply represents the probability of generating that construction
(this probability can be easily calculated based on the generation frequencies
used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}

The vector $\beta$ can be obtained by analyzing the shape of each source of
structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that can be calculated separately):

\begin{equation*}
  \hspace{20pt}
  \beta = \qquad \qquad \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      0 \\ % Text
      0 \\ % Sing
      1 \\ % Tag
      2 \\ % Join
      0 \\ % simplify#1
      2 \\ % simplify#1
      0 \\ % hr
      1 \\ % div
      1 \\ % bold
    };
    \draw[densely dashed] ([xshift=-3pt]v-4-1.south west) -- ([xshift=3pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
  \end{tikzpicture}
  \hspace{20pt}
  \mathcal{P} =
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      p_{Text} \\
      p_{Single} \\
      p_{Tag}  \\
      p_{Join} \\
      p_{simplify\#1} \\
      p_{simplify\#2} \\
      p_{hr} \\
      p_{div} \\
      p_{bold} \\
    };
    \draw[densely dashed] ([xshift=-15pt]v-4-1.south west) -- ([xshift=15pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
  \end{tikzpicture}
\end{equation*}

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized.
%
In this light, \dragenp follows the same approach as \dragen, using our adapted
prediction mechanism to tune the generation probabilities of each each source of
structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
This way she can specify, for instance, a uniform distribution of data
constructors, pattern matching values and function calls or, alternative, a
distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.

@@

It is possible to provide further ``end-to-end'' guarantees on the generated
values if we restrict the generation process to only bare data constructors and
pattern matching values.
%
In particular, we can stop considering pattern matching values as atomic
constructions and start seeing them as compositions of several data
constructors.
%
Then, is possible to obtain the average ``total'' number of data constructors
that our generators will produce, regardless of if they are generated on their
own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$.

The prediction only needs to add the expected number of constructors that we
will additionally generate within pattern matching values to the prediction of
bare data constructors.
%
For instance, we can calculate the total average number of constructors |Text|
and |Join| that we will generate simply by expanding the average number of
generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
corresponding data constructors:
%
{\small
  \begin{align*}
    E^\downarrow\!\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
    E^\downarrow\!\!\left[ Join \right]
    = E\left[ Join \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
  \end{align*}
}
%
Given that each time we generate a value satisfying the first pattern matching
of the function |simplify| we add two |Text| and one |Join| data constructors to
our random value.
%
The case of the second pattern matching of |simplify| follows analogously.
